Computational methods in landscape ecology

### 2024-05-21 Sections/introduction.qmd
Work on intro

Landscape ecology is fun [@Turner1989].
There are many things to explore [@With2019].

>>>>

<!--MH: Flesh-out all paragraphs a little bit adding some details and examples-->

Landscapes are typically defined as mosaics of different land covers, habitats, ecosystems, or land use systems [@Forman1986;@Forman1995] - with empathize on existing heterogeneity in at least one factor of interest [@Turner2015].
Linking the spatial heterogeneity and abiotic as well as biotic processes is the fundamental concept of landscape ecology.
This includes potential feedbacks between heterogeneity and processes, spatial and temporal flows across heterogenous landscapes, or the management of heterogeneity in landscapes [@Risser1984;@Turner1989].
However, due to the spatial context of heterogeneity, the pattern-process link is scale dependent [@Levin1992;@Wiens1989,@Wu2013].

Computational science involves analyzing abstracted core mechanisms of research questions using data and algorithms. 
Thus, computational ecology can be defined if computational science is used to address ecological research questions with focus on complex adaptive systems and data-driven approaches [@Poisot2019].
Computational ecology in general and simulations specifically are important for ecology because ecological data is often context- and scale-dependent and infeasible to study in reproducible, replicable and controlled experimental settings [@Petrovskii2012].

Because landscape ecology is a cross-disciplinary field including [e.g., social sciences, geography, or ecology and evolution @Wiens1997;With2019], the general increase of available data [@Chi2016;@Jaric2020;@Nathan2022;@Wuest2020], or the complex systems characteristics of landscapes and processes [@Newman2019], there is the need for sophisticated computational methods in order to link patterns and processes.

This paper aims to introduce the latest developments of computation methods in landscape ecology.
However, it is not a general introduction into (computational) landscape ecology. 
For a more comprehensive introductions to landscape ecology in general, please see e.g. Turner and Gardner [-@Turner2015] or Gergel and Turner [-@Gergel2017].
